
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Handle Missing Values"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.   
```{r}
library(dplyr)
library(tidyverse)
library(VIM)
df=read_csv("https://bryantstats.github.io/math421/data/adult_census_missing.csv")
colSums(is.na(df))
```

2. Using function `aggr` the VIM package to plot the number of NAs for each column. 

```{r}
aggr(df)
```


3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data.
```{r}
table(df$workclass)
table(df$education)
table(df$marital.status)
table(df$relationship)
table(df$race)
table(df$sex)
table(df$native.country)
```


4. Replace all the forms of missing values found to NA
```{r}
df <- df %>% 
  na_if('Unknown') %>% 
  na_if('Missing') %>% 
  na_if('Not Available') %>%
  na_if('?') %>%
  na_if('Not Known')
```

5. Replot the number of NAs for each column. 
```{r}
aggr(df)
```

6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?
```{r}
df2=drop_na(df)
count(df2)
```

7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}
df2=df %>% 
  fill(age,workclass,fnlwgt,education,education.num,
       marital.status,occupation,relationship,race,sex,
       capital.gain,capital.loss,hours.per.week,native.country,
       income, .direction = 'updown')
count(df2)
```


8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 

```{r}
x <- mean(df$age, na.rm=TRUE)
df$age <- replace_na(df$age, x)

x <- mean(df$fnlwgt, na.rm=TRUE)
df$fnlwgt <- replace_na(df$fnlwgt, x)

x <- mean(df$education.num, na.rm=TRUE)
df$education.num <- replace_na(df$education.num, x)

x <- mean(df$capital.gain, na.rm=TRUE)
df$capital.gain <- replace_na(df$capital.gain, x)

x <- mean(df$capital.loss, na.rm=TRUE)
df$capital.loss <- replace_na(df$capital.loss, x)

x <- mean(df$hours.per.week, na.rm=TRUE)
df$hours.per.week <- replace_na(df$hours.per.week, x)
```





