
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|

```{r}
R = c(0,1,2,3)
A = c(28,34,29,42)
N = c("Tom","Jack","Steve","Ricky")
df = cbind(R,A,N)
df
```

2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).

```{r}
df=read.csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")
```

2. Show the names of the variables in the data
```{r}
names(df)
```

3. How many columns and rows the data have? 
```{r}
ncol(df)
```
```{r}
nrow(df)
```
4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
sum(is.na(df))
```
```{r}
colSums(is.na(df))
```
```{r}
colMeans(is.na(df))
```


5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 

```{r}
names(df)[1] <- 'Date_report'
class(df$Date_report)
```
```{r}
str(df)
```



6. Capitalize the names of all the variables
```{r}
names(df)[1] <- 'DATE_REPORT'
names(df)[2] <- 'STATE'
names(df)[3] <- 'FIPS'
names(df)[4] <- 'CASES'
names(df)[5] <- 'DEATH'
```


7. Find the average number of cases per day. Find the maximum cases a day.  
```{r}
x=aggregate(df$CASES, by=list(DATE_REPORT=df$DATE_REPORT), FUN=mean)
x$DATE_REPORT[x$x==max(x$x)]
```
8. How many states are there in the data?  
```{r}
length(unique(df$STATE))
```


9. Create a new variable `weekdays` to store the weekday for each rows. 

```{r}
df$DATE_REPORT = as.Date(df$DATE_REPORT)
df$WEEK_DAY=weekdays(df$DATE_REPORT)
```

10. Create the categorical variable `death2` variable taking the values as follows
```{r}
DEATH2=c()

for (x in df$DEATH){
  if (x==0){
    DEATH2=append(DEATH2,"no_death")
  }
  else{
    DEATH2=append(DEATH2,"death")
  }
}
df$DEATH2=DEATH2
```

- `has_death` if there is a death that day
-  `no_death` if there is no death that day

Find the frequency and relative frequency of `no_death` and `has_death`. 

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)

```{r}
summary(df$DEATH)
```


12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3

```{r}
DEATH3=c()
a=quantile(df$DEATH)[2]
b=quantile(df$DEATH)[4]

for (x in df$DEATH){
  if (x<a){
    y="low_death"
  }
  else if (x<b){
    y="mid_death"
  }
  else {
    y="high_death"
  }
  DEATH3=append(DEATH3,y)
}
df$DEATH3=DEATH3
```


13. Find the average cases in Rhode Island in 2021

```{r}
df$YEAR <- as.numeric(format(df$DATE_REPORT,'%Y'))
mean(df$CASES[df$STATE=="Rhode Island"&df$YEAR=="2021"])
```


14. Find the median cases by weekdays in Rhode Island in 2021 

```{r}
x=subset(df,df$STATE=="Rhode Island"&df$YEAR==2021)
y=aggregate(x$CASES, by=list(WEEK_DAY=x$WEEK_DAY), FUN=median)
y
```


15. Compare the median cases in Rhode Island in June, July, August and September in 2021.

```{r}
df$MONTH <- as.numeric(format(df$DATE_REPORT,'%m'))
x=subset(df,df$STATE=="Rhode Island"&df$YEAR==2021)
y=aggregate(x$CASES, by=list(MONTH=x$MONTH), FUN=median)
y[6:9,]
```


16.  Find your own dataset, import it and implement the following functions on the data

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when
```{r}
df=read.csv("C:\\Users\\Farida Yesmin\\Desktop\\M421 RStats\\aroshenoor.github.io\\titanic.csv")
head(df)
```
```{r}
str(df)
```

```{r}
names(df)
```
```{r}
mean(df$Age)
```
```{r}
min(df$Age)
```
```{r}
max(df$Age)
```
```{r}
table(df$Age)
```
```{r}
is.na(df$Age)
```
```{r}
colSums(x = cbind(df$Age,df$Pclass,df$Fare),na.rm = TRUE,dims = 1)
```
```{r}
class(df)
```
```{r}
cor(cbind(df$Age,df$Fare,df$Pclass))
```

```{r}
by(df$Fare,list(df$Pclass),mean)
```


17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 

- Create a categorical variable from a continuous variable
- From one categorical variable, create a new  categorical variable with fewer categories

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
